---
title: "Human Activity Recognition"
author: "Nithen Varghese AC"
output: html_document

---
## Abstract
We attempt to quantify the physical activity of partcipant wearing fitness trackers using RandomForest algorithm

##Enable Parallel Processing.
Our implementation of randomForest algorithm is compute intensive due to large number of observations. So we can speed up processing using doParallel package

```{r}
library(doParallel)
cl <- makeCluster(detectCores())
registerDoParallel(cl)

```

## Read input files

we read the input files, the training file  "pml-training.csv" and testing file "pml-testing.csv". The input files contain NA,blanks,Nan and Div/0 all of which will impact the model. So these are converted to NA

```{r}

setwd("F:/ML/workspace/PML")
rawtrain <-read.csv("pml-training.csv",na.strings=c("NA","NaN","","#DIV/0!"))
rawtest <-read.csv("pml-testing.csv",na.strings=c("NA","NaN","","#DIV/0!"))
```

Remove columns that are not meaningful for the current analysis like usernames, timestamps and window details

```{r}
train<- rawtrain[,c(-1,-2,-3,-4,-5,-6,-7)]
test <-rawtest[,c(-1,-2,-3,-4,-5,-6,-7)]

```

# Remove Columns with NA's

Columns with NA's can impact the model creation time and also reduce the accuracy of the created model. 
```{r}
removeNA <- apply(train, 2, function(x) {  sum(is.na(x))})
train <- train[, which(removeNA == 0)]

```

# Partition Training data

We can split the training data into training and test sets to test the model we created.

```{r}
library(caret)

trainingIndex  <- createDataPartition(train$classe, p=.80, list=FALSE)
train.train <- train[ trainingIndex,]
train.test  <- train[-trainingIndex,]

```

#Model Creation

Using the caret package train function, we can create a model using randomForest method. 

```{r}
rf_fit <- train(classe ~ .,train.train,method="rf",tuneGrid=data.frame(mtry=3),trControl=trainControl(method="cv"))


```



# Confusion Matrix
Using the train.test dataset, we can test our model.
```{r}
confusionMatrix(train.test$classe,predict(rf_fit,train.test))

```
From the summary we can see the accuracy of the model and Kappa score's.

Out of sample error rate is denoted by kappa score of 0.9948

#Variable importance

We can find the importance of the variables in determining the model.

```{r}
varImp(rf_fit)

```


# Predicting the test 

We can predict the classes of the provided test set by the below command. 
```{r}
predictin <-predict(rf_fit,newdata=test)

```

